Estimating the genome-wide contribution of selection to temporal allele frequency change

### 2020-05-05 manuscript/manuscript.tex

Finally we used forward-in-time simulations to explore the conditions under
which temporal and convergent correlations arise. We show a subset of our
results for a model of Gaussian stabilizing selection (GSS) on a phenotype
where directional selection is induced by a sudden shift in the optimum
phenotype of varying magnitudes (Figure \ref{fig:figure-4}A).  We find that
positive temporal covariances are produced by such selection (Figure
\ref{fig:figure-4}B), and that these positive temporal covariances can compound
together to generate a large proportion of allele frequency change being due to
selection (i.e. large G) over the relatively short time periods similar to our
analyzed selection datasets span (Figure \ref{fig:figure-4}C). The magnitude of
these covariances increases with the strength of selection, i.e. the variance
in fitness, such that stronger selection generates larger proportions of allele
frequency change.  We find a similar picture of stronger convergent selection
pressures generating larger convergence correlations.

These results are often relatively insensitive to the number of loci underlying
the trait, suggesting that they may be capturing highly polygenic signals.
Indeed if only a small number of loci influence the trait, the $G(t)$
trajectories are typically much more stochastic, such that replicates do not
reliably generate temporal covariances (compare Figure \ref{fig:figure-1}B to
Supplementary Material Figure \ref{suppfig:sim-expfit-covs}).

Additionally, we explored other selection schemes. We find that the long term
dynamics of the covariances under truncation selection, which generates
substantial epistasis, are richer than we see under GSS and multiplicative
selection (Supplementary Material Figure \ref{suppfig:supp-trunc}).  We also
conducted simulations of purifying selection alone (i.e. background selection)
and find that this can also generate positive temporal covariances
(Supplementary Material Figure \ref{suppfig:supp-bgs-covs}), thus it is
unlikely that all of the allele frequency change we see is the result of the
novel selection pressure the populations are exposed to, and some of this
selection may be ongoing purifying selection. Only in the case of the
Longshanks experiment, does the control line mean that we can quantify
selection that is due to the novel selection pressure.

While none of our experiments have selected the populations in divergent
directions, in our simulations we find that such selection can generate
negative convergent correlations. This suggests that selection experiments
combining multiple replicates, control lines, as well as divergent selection
pressures might be quite informative for unpacking the contribution of
particular selection pressures to genome-wide allele frequency change.

>>>>

Finally we used forward-in-time simulations to explore the conditions under
which temporal and convergent correlations arise. We show a subset of our
results for a model of Gaussian stabilizing selection (GSS) on a
phenotype  \gc{DO WE actually need to say Gaussian here? Or just in
  methods. }
where directional selection is induced by a sudden shift in the optimum
phenotype of varying magnitudes (Figure \ref{fig:figure-4}A).  We find that
positive temporal covariances are produced by such selection (Figure
\ref{fig:figure-4}B), and that these positive temporal covariances can compound
together to generate a large proportion of allele frequency change being due to
selection (i.e. large G) over the relatively short time periods similar to our
analyzed selection datasets span (Figure \ref{fig:figure-4}C). The magnitude of
these covariances increases with the strength of selection, i.e. the variance
in fitness, such that stronger selection generates larger proportions of allele
frequency change.  We find a similar picture of stronger convergent selection
pressures generating larger convergence correlations \gc{(Figure {fig:figure-4}D)}.

These results are often relatively insensitive to the number of loci underlying
the trait, suggesting that they may be capturing highly polygenic signals.
Indeed if only a small number of loci influence the trait, the $G(t)$
trajectories are typically much more stochastic, such that replicates do not
reliably generate temporal covariances (compare Figure
\ref{fig:figure-1}B to Supplementary Material Figure \ref{suppfig:sim-expfit-covs})
\gc{This again suggests that
  the genome-wide linked selection response we see in the empirical data is highly
polygenic. Using our simulations we find that sampling only every 10
generations does indeed mean that our estimates of $G(t)$ are an
underestimate of the contribution of linked selection as they cannot include the covariance between closely
spaced generations (see XXXX).
}

Additionally, we explored other selection schemes. We find that the long term
dynamics of the covariances under \gc{directional} truncation selection, which generates
substantial epistasis, are richer than we see under GSS and multiplicative
selection (Supplementary Material Figure \ref{suppfig:supp-trunc}).  We also
conducted simulations of purifying selection alone (i.e. background selection)
and find that this can also generate positive temporal covariances
(Supplementary Material Figure \ref{suppfig:supp-bgs-covs}), thus it is
unlikely that all of the allele frequency change we see is the result of the
novel selection pressure the populations are exposed to, and some of this
selection may be ongoing purifying selection. Only in the case of the
Longshanks experiment, does the control line mean that we can quantify
selection that is due to the novel selection pressure.

While none of our experiments have selected the populations in divergent
directions, in our simulations we find that such selection can generate
negative convergent correlations \gc{(Figure {fig:figure-4}D)}. This suggests that selection experiments
combining multiple replicates, control lines, as well as divergent selection
pressures might be quite informative for unpacking the contribution of
particular selection pressures to genome-wide allele frequency change.
